---
title: "Additional File 5"
subtitle: "Sensitivity analysis of nonlinear associations by exclusion of single lifestyle components from the Healthy Lifestyle Index"

#author: "Sairah L.F. Chen"
#date: "14 12 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(rms)
library(survival)
library(tidyverse)
library(ggpubr)
library(cowplot)

casesIncluded <- read.csv2(file="../data/casesIncluded25112020.csv")
```

```{r preparing reduced HLI vars, include=FALSE}

casesIncluded$HLInoSmoking <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoPA <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$smokingScore +
  casesIncluded$BMIScore

casesIncluded$HLInoBMI <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$smokingScore

casesIncluded$HLInoDiet<- casesIncluded$smokingScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoAlcohol <- casesIncluded$dietScore +
  casesIncluded$smokingScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore


```


```{r, setup datadist, include=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")
ddist$limits["Adjust to", "HLIScore"] <- 5
```

## Postmenopausal breast cancer 

<span style="color:gray"> Nonlinear associations of HLI score excluding each lifestyle component one-by-one and postmenopausal breast cancer incidence. Shaded region visualises 95% CI. Restricted cubic splines fitted with three knots for each  model </span>


```{r, echo=FALSE, warning=FALSE}


breastNoSmoking <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoSmoking <- ggplot(rms:::Predict(breastNoSmoking, HLInoSmoking=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Breast cancer without smoking") +
  theme(legend.position = "none")

plotbreastNoSmoking


breastNoPA <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoPA <- ggplot(rms:::Predict(breastNoPA, HLInoPA=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Breast cancer without PA") +
  theme(legend.position = "none")

plotbreastNoPA


breastNoBMI <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoBMI <- ggplot(rms:::Predict(breastNoBMI, HLInoBMI=seq(1, 16),
                                        ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Breast cancer without BMI") +
  theme(legend.position = "none")

plotbreastNoBMI


breastNoDiet <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoDiet <- ggplot(rms:::Predict(breastNoDiet, HLInoDiet=seq(1, 16),
                                         ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Breast cancer without diet") +
  theme(legend.position = "none")

plotbreastNoDiet


breastNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoAlcohol <- ggplot(rms:::Predict(breastNoAlcohol, HLInoAlcohol=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Breast cancer without alcohol") +
  theme(legend.position = "none")

plotbreastNoAlcohol


```

## Lung cancer 

<span style="color:gray"> Nonlinear associations of HLI score excluding each lifestyle component one-by-one and lung cancer incidence. Shaded region visualises 95% CI. Restricted cubic splines fitted with five knots for each  model </span>

```{r, echo=FALSE, warning=FALSE}
lungNoSmoking <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoSmoking <- ggplot(rms:::Predict(lungNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer without smoking") +
  theme(legend.position = "none")

plotLungNoSmoking

lungNoPA <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoPA <- ggplot(rms:::Predict(lungNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer without PA") +
  theme(legend.position = "none")

plotLungNoPA

lungNoBMI <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoBMI <- ggplot(rms:::Predict(lungNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer without BMI") +
  theme(legend.position = "none")

plotLungNoBMI


lungNoDiet <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoDiet <- ggplot(rms:::Predict(lungNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer without diet") +
  theme(legend.position = "none")

plotLungNoDiet


lungNoAlcohol <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoAlcohol <- ggplot(rms:::Predict(lungNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer without alcohol") +
  theme(legend.position = "none")

plotLungNoAlcohol


```


## Postmenopausal endometrial cancer 

<span style="color:gray"> Nonlinear associations of HLI score excluding each lifestyle component one-by-one and postmenopausal endometrial cancer incidence. Shaded region visualises 95% CI. Restricted cubic splines fitted with three knots for each  model </span>


```{r, echo=FALSE, warning=FALSE}


endometrialNoSmoking <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoSmoking <- ggplot(rms:::Predict(endometrialNoSmoking, HLInoSmoking=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Endometrial cancer without smoking") +
  theme(legend.position = "none")

plotEndometrialNoSmoking


endometrialNoPA <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoPA <- ggplot(rms:::Predict(endometrialNoPA, HLInoPA=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Endometrial cancer without PA") +
  theme(legend.position = "none")

plotEndometrialNoPA


endometrialNoBMI <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoBMI <- ggplot(rms:::Predict(endometrialNoBMI, HLInoBMI=seq(1, 16),
                                             ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Endometrial cancer without BMI") +
  theme(legend.position = "none")

plotEndometrialNoBMI


endometrialNoDiet <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoDiet <- ggplot(rms:::Predict(endometrialNoDiet, HLInoDiet=seq(1, 16),
                                              ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Endometrial cancer without diet") +
  theme(legend.position = "none")

plotEndometrialNoDiet


endometrialNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoAlcohol <- ggplot(rms:::Predict(endometrialNoAlcohol, HLInoAlcohol=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score without alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Endometrial cancer without alcohol") +
  theme(legend.position = "none")

plotEndometrialNoAlcohol


```





rmarkdown::render("C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/r_docs/nonlinear_sensitivity.Rmd", output_format= "pdf_document", output_file = "C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/output/nonlinear_sensitivity.pdf")


