---
title: "Additional File 5"
subtitle: "Sensitivity analysis of nonlinear associations by exclusion of single lifestyle components from the Healthy Lifestyle Index (HLI)"

#author: "Sairah L.F. Chen"
#date: "14 12 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(rms)
library(survival)
library(tidyverse)
library(ggpubr)
library(cowplot)

casesIncluded <- read.csv2(file="../data/casesIncluded25112020.csv")
```

This additional file displays results from sensitivity analyses conducted on associations with the HLI modeled as a nonlinear term using restricted cubic splines (RCS) with three knots. Single lifestyle components were excluded from the HLI, creating five reduced HLIs. Thus, five estimated associations were obtained for each cancer outcome. Here, we visualise these results in plots. 


```{r preparing reduced HLI vars, include=FALSE}

casesIncluded$HLInoSmoking <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoPA <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$smokingScore +
  casesIncluded$BMIScore

casesIncluded$HLInoBMI <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$smokingScore

casesIncluded$HLInoDiet<- casesIncluded$smokingScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoAlcohol <- casesIncluded$dietScore +
  casesIncluded$smokingScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore


```


```{r, setup datadist, include=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")
ddist$limits["Adjust to", "HLIScore"] <- 5
```

Figure 1. Postmenopausal breast cancer incidence and its association with HLIs excluding single lifestyle components. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}


breastNoSmoking <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoSmoking <- ggplot(rms:::Predict(breastNoSmoking, HLInoSmoking=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoPA <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoPA <- ggplot(rms:::Predict(breastNoPA, HLInoPA=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding physical activity",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



breastNoBMI <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoBMI <- ggplot(rms:::Predict(breastNoBMI, HLInoBMI=seq(1, 16),
                                        ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoDiet <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoDiet <- ggplot(rms:::Predict(breastNoDiet, HLInoDiet=seq(1, 16),
                                         ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoAlcohol <- ggplot(rms:::Predict(breastNoAlcohol, HLInoAlcohol=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breast <- ggarrange(plotbreastNoSmoking,
                    plotbreastNoPA,
                    plotbreastNoDiet,
                    plotbreastNoBMI,
                    plotbreastNoAlcohol,
                    ncol=2, nrow=3,
                    labels=c("a","b","c","d","e"))
breast

```

\newpage

Figure 2. Colorectal cancer incidence and its association with HLIs excluding single lifestyle components. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r, echo=FALSE, warning=FALSE}
crcNoSmoking <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoSmoking <- ggplot(rms:::Predict(crcNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Colorectal cancer excluding smoking") +
  theme(legend.position = "none")

plotLungNoSmoking

crcNoPA <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoPA <- ggplot(rms:::Predict(crcNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Colorectal cancer excluding PA") +
  theme(legend.position = "none")

plotCrcNoPA

crcNoBMI <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoBMI <- ggplot(rms:::Predict(crcNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Colorectal cancer excluding BMI") +
  theme(legend.position = "none")

plotCrcNoBMI


crcNoDiet <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoDiet <- ggplot(rms:::Predict(crcNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Colorectal cancer excluding diet") +
  theme(legend.position = "none")

plotCrcNoDiet


crcNoAlcohol <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoAlcohol <- ggplot(rms:::Predict(crcNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Colorectal cancer excluding alcohol") +
  theme(legend.position = "none")

plotCrcNoAlcohol

crc <- ggarrange(plotbreastNoSmoking,
                 plotbreastNoPA,
                 plotbreastNoDiet,
                 plotbreastNoBMI,
                 plotbreastNoAlcohol,
                 ncol=2, nrow=3,
                 labels=c("a","b","c","d","e"))
crc


```


Figure 3. Lung cancer incidence and its association with HLIs excluding single lifestyle components. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 

```{r, echo=FALSE, warning=FALSE}
lungNoSmoking <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoSmoking <- ggplot(rms:::Predict(lungNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer excluding smoking") +
  theme(legend.position = "none")

plotLungNoSmoking

lungNoPA <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoPA <- ggplot(rms:::Predict(lungNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer excluding PA") +
  theme(legend.position = "none")

plotLungNoPA

lungNoBMI <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoBMI <- ggplot(rms:::Predict(lungNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer excluding BMI") +
  theme(legend.position = "none")

plotLungNoBMI


lungNoDiet <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoDiet <- ggplot(rms:::Predict(lungNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer excluding diet") +
  theme(legend.position = "none")

plotLungNoDiet


lungNoAlcohol <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoAlcohol <- ggplot(rms:::Predict(lungNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Lung cancer excluding alcohol") +
  theme(legend.position = "none")

plotLungNoAlcohol


```


## Postmenopausal endometrial cancer 



```{r, echo=FALSE, warning=FALSE}


endometrialNoSmoking <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoSmoking <- ggplot(rms:::Predict(endometrialNoSmoking, HLInoSmoking=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Postmenopausal endometrial cancer excluding smoking") +
  theme(legend.position = "none")

plotEndometrialNoSmoking


endometrialNoPA <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoPA <- ggplot(rms:::Predict(endometrialNoPA, HLInoPA=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Postmenopausal endometrial cancer excluding PA") +
  theme(legend.position = "none")

plotEndometrialNoPA


endometrialNoBMI <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoBMI <- ggplot(rms:::Predict(endometrialNoBMI, HLInoBMI=seq(1, 16),
                                             ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Postmenopausal endometrial cancer excluding BMI") +
  theme(legend.position = "none")

plotEndometrialNoBMI


endometrialNoDiet <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoDiet <- ggplot(rms:::Predict(endometrialNoDiet, HLInoDiet=seq(1, 16),
                                              ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Postmenopausal endometrial cancer excluding diet") +
  theme(legend.position = "none")

plotEndometrialNoDiet


endometrialNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoAlcohol <- ggplot(rms:::Predict(endometrialNoAlcohol, HLInoAlcohol=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  ggtitle("Postmenopausal endometrial cancer excluding alcohol") +
  theme(legend.position = "none")

plotEndometrialNoAlcohol


```





rmarkdown::render("C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/r_docs/nonlinear_sensitivity.Rmd", output_format= "pdf_document", output_file = "C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/output/nonlinear_sensitivity.pdf")


