---
title: "Additional File 5"
subtitle: "Sensitivity analysis of nonlinear associations by exclusion of single lifestyle components from the Healthy Lifestyle Index (HLI)"

#author: "Sairah L.F. Chen"
#date: "14 12 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(rms)
library(survival)
library(tidyverse)
library(ggpubr)
library(cowplot)

casesIncluded <- read.csv2(file="../data/casesIncluded25112020.csv")
```

Note: could not get these models to have their reference to be set to HLI=5 like in the main analysis. Will work on it. \
\
\

This additional file displays results from sensitivity analyses conducted on associations with the HLI modeled as a nonlinear term using restricted cubic splines (RCS) with three knots. Single lifestyle components were excluded from the HLI, creating five reduced HLIs. Thus, five estimated associations were obtained for each cancer outcome. Here, we visualise these results in plots. 


```{r preparing reduced HLI vars, include=FALSE}

casesIncluded$HLInoSmoking <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoPA <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$smokingScore +
  casesIncluded$BMIScore

casesIncluded$HLInoBMI <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$smokingScore

casesIncluded$HLInoDiet<- casesIncluded$smokingScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoAlcohol <- casesIncluded$dietScore +
  casesIncluded$smokingScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore


```


```{r, setup datadist, include=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")
ddist$limits["Adjust to", "HLIScore"] <- 5

```


\newpage

Figure 1. Postmenopausal breast cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education, hormone replacement therapy status, oral contraceptive ever use, parity, age at menarche, breastfeeding, and family history of breast cancer in the first degree. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r postmenopausal breast, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}


breastNoSmoking <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoSmoking <- ggplot(rms:::Predict(breastNoSmoking, HLInoSmoking=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoPA <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoPA <- ggplot(rms:::Predict(breastNoPA, HLInoPA=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding physical activity",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



breastNoBMI <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoBMI <- ggplot(rms:::Predict(breastNoBMI, HLInoBMI=seq(1, 16),
                                        ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoDiet <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoDiet <- ggplot(rms:::Predict(breastNoDiet, HLInoDiet=seq(1, 16),
                                         ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breastNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoAlcohol <- ggplot(rms:::Predict(breastNoAlcohol, HLInoAlcohol=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




breast <- ggarrange(plotbreastNoSmoking,
                    plotbreastNoPA,
                    plotbreastNoDiet,
                    plotbreastNoBMI,
                    plotbreastNoAlcohol,
                    ncol=2, nrow=3,
                    labels=c("a","b","c","d","e"))
breast

```

\newpage

Figure 2. Colorectal cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r colorectal, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}
crcNoSmoking <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoSmoking <- ggplot(rms:::Predict(crcNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


crcNoPA <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoPA <- ggplot(rms:::Predict(crcNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


crcNoBMI <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoBMI <- ggplot(rms:::Predict(crcNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




crcNoDiet <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoDiet <- ggplot(rms:::Predict(crcNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




crcNoAlcohol <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotCrcNoAlcohol <- ggplot(rms:::Predict(crcNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



crc <- ggarrange(plotCrcNoSmoking,
                 plotCrcNoPA,
                 plotCrcNoDiet,
                 plotCrcNoBMI,
                 plotCrcNoAlcohol,
                 ncol=2, nrow=3,
                 labels=c("a","b","c","d","e"))
crc


```

\newpage

Figure 3. Lung cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 

```{r lung, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}
lungNoSmoking <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoSmoking <- ggplot(rms:::Predict(lungNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


lungNoPA <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoPA <- ggplot(rms:::Predict(lungNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



lungNoBMI <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoBMI <- ggplot(rms:::Predict(lungNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




lungNoDiet <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoDiet <- ggplot(rms:::Predict(lungNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




lungNoAlcohol <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoAlcohol <- ggplot(rms:::Predict(lungNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




lung <- ggarrange(plotLungNoSmoking,
                  plotLungNoPA,
                  plotLungNoDiet,
                  plotLungNoBMI,
                  plotLungNoAlcohol,
                 ncol=2, nrow=3,
                 labels=c("a","b","c","d","e"))
lung


```

\newpage

Figure 4. Postmenopausal endometrial cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education, hormone replacement therapy status, oral contraceptive ever use, parity, age at menarche, and breastfeeding. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r postmenopausal endometrial, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}


endometrialNoSmoking <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoSmoking <- ggplot(rms:::Predict(endometrialNoSmoking, HLInoSmoking=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




endometrialNoPA <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoPA <- ggplot(rms:::Predict(endometrialNoPA, HLInoPA=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




endometrialNoBMI <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoBMI <- ggplot(rms:::Predict(endometrialNoBMI, HLInoBMI=seq(1, 16),
                                             ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




endometrialNoDiet <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoDiet <- ggplot(rms:::Predict(endometrialNoDiet, HLInoDiet=seq(1, 16),
                                              ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




endometrialNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrialNoAlcohol <- ggplot(rms:::Predict(endometrialNoAlcohol, HLInoAlcohol=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")





endometrial <- ggarrange(plotEndometrialNoSmoking,
                        plotEndometrialNoPA,
                        plotEndometrialNoDiet,
                        plotEndometrialNoBMI,
                        plotEndometrialNoAlcohol,
                        ncol=2, nrow=3,
                        labels=c("a","b","c","d","e"))
endometrial


```

\newpage

Figure 5. Postmenopausal ovarian cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education, hormone replacement therapy status, oral contraceptive ever use, parity, age at menarche, and breastfeeding.  The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 




```{r postmenopausal ovarian, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}


ovarianNoSmoking <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarianNoSmoking <- ggplot(rms:::Predict(ovarianNoSmoking, HLInoSmoking=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




ovarianNoPA <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarianNoPA <- ggplot(rms:::Predict(ovarianNoPA, HLInoPA=seq(1, 16),
                                            ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




ovarianNoBMI <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarianNoBMI <- ggplot(rms:::Predict(ovarianNoBMI, HLInoBMI=seq(1, 16),
                                             ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




ovarianNoDiet <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarianNoDiet <- ggplot(rms:::Predict(ovarianNoDiet, HLInoDiet=seq(1, 16),
                                              ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




ovarianNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarianNoAlcohol <- ggplot(rms:::Predict(ovarianNoAlcohol, HLInoAlcohol=seq(1, 16),
                                                 ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")





ovarian <- ggarrange(plotOvarianNoSmoking,
                        plotOvarianNoPA,
                        plotOvarianNoDiet,
                        plotOvarianNoBMI,
                        plotOvarianNoAlcohol,
                        ncol=2, nrow=3,
                        labels=c("a","b","c","d","e"))
ovarian


```

\newpage


Figure 6. Pancreatic cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r pancreatic, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}
pancreaticNoSmoking <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreaticNoSmoking <- ggplot(rms:::Predict(pancreaticNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


pancreaticNoPA <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreaticNoPA <- ggplot(rms:::Predict(pancreaticNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


pancreaticNoBMI <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreaticNoBMI <- ggplot(rms:::Predict(pancreaticNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




pancreaticNoDiet <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreaticNoDiet <- ggplot(rms:::Predict(pancreaticNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




pancreaticNoAlcohol <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreaticNoAlcohol <- ggplot(rms:::Predict(pancreaticNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



pancreatic <- ggarrange(plotPancreaticNoSmoking,
                 plotPancreaticNoPA,
                 plotPancreaticNoDiet,
                 plotPancreaticNoBMI,
                 plotPancreaticNoAlcohol,
                 ncol=2, nrow=3,
                 labels=c("a","b","c","d","e"))
pancreatic

```




\newpage


Figure 7. Kidney cancer incidence and its association with HLIs excluding single lifestyle components. Models are adjusted for education. The shaded regions represent the 95% confidence intervals and the reference HLI score is set to 5. 



```{r kidney, echo=FALSE, warning=FALSE, fig.height=10, fig.width=10}
kidneyNoSmoking <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotKidneyNoSmoking <- ggplot(rms:::Predict(kidneyNoSmoking, HLInoSmoking=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


kidneyNoPA <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotKidneyNoPA <- ggplot(rms:::Predict(kidneyNoPA, HLInoPA=seq(1, 16),
                                     ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")


kidneyNoBMI <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotKidneyNoBMI <- ggplot(rms:::Predict(kidneyNoBMI, HLInoBMI=seq(1, 16),
                                      ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




kidneyNoDiet <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotKidneyNoDiet <- ggplot(rms:::Predict(kidneyNoDiet, HLInoDiet=seq(1, 16),
                                       ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")




kidneyNoAlcohol <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotKidneyNoAlcohol <- ggplot(rms:::Predict(kidneyNoAlcohol, HLInoAlcohol=seq(1, 16),
                                          ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score excluding alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
  theme(legend.position = "none")



kidney <- ggarrange(plotKidneyNoSmoking,
                 plotKidneyNoPA,
                 plotKidneyNoDiet,
                 plotKidneyNoBMI,
                 plotKidneyNoAlcohol,
                 ncol=2, nrow=3,
                 labels=c("a","b","c","d","e"))
kidney

```



rmarkdown::render("C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/r_docs/nonlinear_sensitivity.Rmd", output_format= "pdf_document", output_file = "C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/output/nonlinear_sensitivity.pdf")