---
title: "Additional File 3"
subtitle: "Demonstrating model fit of associations between the HLI modeled as a nonlinear term and incidences of common cancers in NOWAC"

#author: "Sairah L.F. Chen"
#date: "08/07/2020"
output: pdf_document


---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rms)
library(survival)
library(tidyverse)
library(ggpubr)
library(cowplot)


casesIncluded <- read.csv2(file="../data/casesIncluded25112020.csv")
```

This additional document displays Akaike's information criteria (AIC) values and plots generated for Cox proportional hazards regression models to estimate the associations between the HLI modeled as a nonlinear term and incidences of common cancers.  Restricted cubic splines (RCS) were employed to model the HLI as a nonlinear term, where terms modeled with three, four, and five knots located at the percentiles have been tested. AIC values and plots of linear associations are also included for comparison.


```{r, setup datadist, include=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")
ddist$limits["Adjust to", "HLIScore"] <- 5
```

```{r best fit plots , include=FALSE}


# breast, 3 knots
coxspline3Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)



plotBreast3 <- ggplot(rms:::Predict(coxspline3Breast, HLIScore=seq(1, 20), 
                                    ref.zero = TRUE,
                                    fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) +
              geom_hline(yintercept=1, linetype="dashed") +
              ggtitle("Breast cancer") +
              theme(legend.position = "none")

plotBreast3

#crc, 3 knots

coxspline3Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotColorectal3 <- ggplot(rms:::Predict(coxspline3Colorectal, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("Colorectal cancer")

plotColorectal3


# lung, 3 knots

coxspline3Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotLung3 <- ggplot(rms:::Predict(coxspline3Lung, HLIScore=seq(1, 20),
                                  ref.zero = TRUE,
                                  fun=exp),
                          xlab= "HLI Score",
                          ylab="HR",
                          adj.subtitle = FALSE) + 
              geom_hline(yintercept=1, linetype="dashed") +
              ggtitle("3 knots")

plotLung3

# endometrial, 3 knots
coxspline3Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrial3 <- ggplot(rms:::Predict(coxspline3Endometrial, HLIScore=seq(1, 20),
                                         ref.zero = TRUE,
                                         fun=exp),
                        xlab= "HLI Score",
                        ylab="HR",
                        adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("Endometrial cancer")

plotEndometrial3


# ovarian, 3 knots
coxspline3Ovarian <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLIScore, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

plotOvarian3 <- ggplot(rms:::Predict(coxspline3Ovarian, HLIScore=seq(1, 20),
                                     ref.zero = TRUE,
                                     fun=exp),
                        xlab= "HLI Score",
                        ylab="HR",
                        adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("Ovarian cancer")

plotOvarian3 

# pancreatic, 3 knots
coxspline3Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreatic3 <- ggplot(rms:::Predict(coxspline3Pancreatic, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                        xlab= "HLI Score",
                        ylab="HR",
                        adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("Pancreatic cancer")

plotPancreatic3

# kidney, 3 knots

coxspline3Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotKidney3 <- ggplot(rms:::Predict(coxspline3Kidney, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("Kidney cancer")

plotKidney3

#png("rcs3Knots.png", width= 1600, height = 1600)
#rcs3Knots <- plot_grid(plotBreast3,
#             plotColorectal3, 
#             plotLung3,
#             plotEndometrial3,
#             plotOvarian3,
#             plotPancreatic3,
#             plotKidney3,
#             ncol=3, nrow=3)
#dev.off()
#rcs3Knots

#rcs3KnotsLabeled <- add_sub(rcs3Knots, 
#                                       "Figure 1. Non-linear association of HLI score and risk for common cancers. \nShaded region visualises 95% CI. Restricted cubic splines fitted with three \nknots set at HLI score 1, 13, and 20 for each cancer model", 
#                            size=12,
#                            x = 0,
#                            h = 0)
#ggdraw(rcs3KnotsLabeled)






flexibleBestFit <- plot_grid(plotBreast3,
             plotColorectal3, 
             plotLung3,
             plotEndometrial3,
             plotOvarian3,
             plotPancreatic3,
             plotKidney3,
             ncol=3, nrow=3) 
flexibleBestFit



```



```{r, echo=FALSE, warning=FALSE}



coxspline5Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 5) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 4) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxBreast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ HLIScore + 
                   education +
                   HRTStatus +
                   OCEverUse +
                   parityCat +
                   familyHistBC +
                   ageMenarche +
                   breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline5Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxColorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline5Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxLung <- cph(Surv(ageEntry, ageExit, statusLung) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)




coxspline5Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 5) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 4) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)



coxEndometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ HLIScore + 
                   education +
                   HRTStatus +
                   OCEverUse +
                   parityCat +
                   ageMenarche +
                   breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)
coxspline5Ovarian <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLIScore, 5) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Ovarian <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ rcs(HLIScore, 4) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)



coxOvarian <- cph(Surv(ageMenopause, ageExit, statusOvarian) ~ HLIScore + 
                   education +
                   HRTStatus +
                   OCEverUse +
                   parityCat +
                   ageMenarche +
                   breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)


coxspline5Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxPancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)


coxspline5Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)


coxspline4Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxKidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)






AIC <- cbind(c(AIC(coxspline5Breast),
               AIC(coxspline4Breast),
               AIC(coxspline3Breast),
               AIC(coxBreast)),
             c(AIC(coxspline5Colorectal),
               AIC(coxspline4Colorectal),
               AIC(coxspline3Colorectal),
               AIC(coxColorectal)),
             c(AIC(coxspline5Lung),
               AIC(coxspline4Lung),
               AIC(coxspline3Lung),
               AIC(coxLung)),
             c(AIC(coxspline5Endometrial),
               AIC(coxspline4Endometrial),
               AIC(coxspline3Endometrial),
               AIC(coxEndometrial)),
             c(AIC(coxspline5Ovarian),
               AIC(coxspline4Ovarian),
               AIC(coxspline3Ovarian),
               AIC(coxOvarian)),
             c(AIC(coxspline5Pancreatic),
               AIC(coxspline4Pancreatic),
               AIC(coxspline3Pancreatic),
               AIC(coxPancreatic)),
             c(AIC(coxspline5Kidney), 
               AIC(coxspline4Kidney), 
               AIC(coxspline3Kidney), 
               AIC(coxKidney)))
rownames(AIC) <- c("5 knots",
                   "4 knots",
                   "3 knots",
                   "Linear")
colnames(AIC) <- c("Breast",
                   "Colorectal",
                   "Lung",
                   "Endometrial",
                   "Ovarian",
                   "Pancreatic",
                   "Kidney")

knitr::kable(as.data.frame(AIC), digits = 2,
             caption = "AIC values for linear and nonlinear models (3-5 RCS knots) among common cancers")
```


\newpage

All of the following plots use the HLI score of 5 as the reference. Hazard ratios (HRs) are present on the vertical axes and HLI scores are present on the horizontal axes. Shaded regions represent 95% confidence intervals.\
\
\

Postmenopausal breast cancer

```{r breast knot plots, echo=FALSE}


# breast, 5 knots


plotBreast5 <- ggplot(rms:::Predict(coxspline5Breast, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("5 knots")



# breast, 4 knots



plotBreast4 <- ggplot(rms:::Predict(coxspline4Breast, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("4 knots")




#breast, Linear



plotBreast <- ggplot(rms:::Predict(coxBreast, HLIScore=seq(1, 20),
                                   ref.zero = TRUE,
                                    fun=exp),
                            xlab= "HLI Score",
                            ylab="HR",
                            adj.subtitle = FALSE) + 
              geom_hline(yintercept=1, linetype="dashed") +
              ggtitle("Linear")



```


```{r, breast fig, echo=FALSE}
breastSplineOptions <- ggarrange(plotBreast5,
                               plotBreast4,
                               plotBreast,
                               ncol=2, nrow=2,
                               labels=c("a","b","c"))

#annotate_figure(breastSplineOptions, top= text_grob("Postmenopausal breast cancer", windowsFonts(Times=windowsFont("TT Times New Roman"))))


breastSplineOptions
```

\newpage

Colorectal cancer



```{r crc knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# crc, 5 knots


plotColorectal5 <- ggplot(rms:::Predict(coxspline5Colorectal, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("5 knots")



# crc, 4 knots



plotColorectal4 <- ggplot(rms:::Predict(coxspline4Colorectal, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("4 knots")



#crc, Linear



plotColorectal <- ggplot(rms:::Predict(coxColorectal, HLIScore=seq(1, 20),
                                       ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                  geom_hline(yintercept=1, linetype="dashed") +
                  ggtitle("Linear")



colorectalSplineOptions <- ggarrange(plotColorectal5,
                               plotColorectal4,
                               plotColorectal,
                               ncol=2, nrow=2,
                               labels=c("a","b","c"))

colorectalSplineOptions

```

\newpage

Lung cancer



```{r lung knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")



#lung, 5 knots




plotLung5 <- ggplot(rms:::Predict(coxspline5Lung, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                                    xlab= "HLI Score",
                                    ylab="HR",
                                    adj.subtitle = FALSE) + 
              geom_hline(yintercept=1, linetype="dashed") +
              ggtitle("Lung cancer")




# lung, 4 knots



plotLung4 <- ggplot(rms:::Predict(coxspline4Lung, HLIScore=seq(1, 20),
                                  ref.zero = TRUE,
                                  fun=exp),
                          xlab= "HLI Score",
                          ylab="HR",
                          adj.subtitle = FALSE) + 
              geom_hline(yintercept=1, linetype="dashed") +
              ggtitle("4 knots")



#lung, Linear



plotLung <- ggplot(rms:::Predict(coxLung, HLIScore=seq(1, 20),
                                 ref.zero = TRUE,
                                 fun=exp),
                          xlab= "HLI Score",
                          ylab="HR",
                          adj.subtitle = FALSE) + 
            geom_hline(yintercept=1, linetype="dashed") +
            ggtitle("Linear")



lungSplineOptions <- ggarrange(plotLung3,
                               plotLung4,
                               plotLung,
                               ncol=2, nrow=2,
                               labels=c("a","b","c"))

lungSplineOptions

```

\newpage

Postmenopausal endometrial cancer



```{r endometrial knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# endo, 5 knots


plotEndometrial5 <- ggplot(rms:::Predict(coxspline5Endometrial, HLIScore=seq(1, 20),
                                         ref.zero = TRUE,
                                         fun=exp),
                                  xlab= "HLI Score",
                                  ylab="HR",
                                  adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("5 knots")



# endo, 4 knots



plotEndometrial4 <- ggplot(rms:::Predict(coxspline4Endometrial, HLIScore=seq(1, 20),
                                         ref.zero = TRUE,
                                         fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("4 knots")



#endo, Linear



plotEndometrial <- ggplot(rms:::Predict(coxEndometrial, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
    ggtitle("Linear")



endometrialSplineOptions <- ggarrange(plotEndometrial5,
                                      plotEndometrial4,
                                      plotEndometrial,
                                      ncol=2, nrow=2,
                                      labels=c("a","b","c"))

endometrialSplineOptions

```




\newpage

Postmenopausal ovarian cancer



```{r ovarian knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# endo, 5 knots


plotOvarian5 <- ggplot(rms:::Predict(coxspline5Ovarian, HLIScore=seq(1, 20),
                                         ref.zero = TRUE,
                                         fun=exp),
                                  xlab= "HLI Score",
                                  ylab="HR",
                                  adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("5 knots")



# endo, 4 knots



plotOvarian4 <- ggplot(rms:::Predict(coxspline4Ovarian, HLIScore=seq(1, 20),
                                         ref.zero = TRUE,
                                         fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("4 knots")



#endo, Linear



plotOvarian <- ggplot(rms:::Predict(coxOvarian, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
    geom_hline(yintercept=1, linetype="dashed") +
    ggtitle("Linear")



ovarianSplineOptions <- ggarrange(plotOvarian5,
                                      plotOvarian4,
                                      plotOvarian,
                                      ncol=2, nrow=2,
                                      labels=c("a","b","c"))

ovarianSplineOptions

```


\newpage

Pancreatic cancer



```{r pancreatic knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# pancreatic, 5 knots


plotPancreatic5 <- ggplot(rms:::Predict(coxspline5Pancreatic, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                                xlab= "HLI Score",
                                ylab="HR",
                                adj.subtitle = FALSE) + 
                  geom_hline(yintercept=1, linetype="dashed") +
                  ggtitle("5 knots")



#pancreatic, 4 knots



plotPancreatic4 <- ggplot(rms:::Predict(coxspline4Pancreatic, HLIScore=seq(1, 20),
                                        ref.zero = TRUE,
                                        fun=exp),
                          xlab= "HLI Score",
                          ylab="HR",
                          adj.subtitle = FALSE) + 
                    geom_hline(yintercept=1, linetype="dashed") +
                    ggtitle("4 knots")



#pancreatic, Linear



plotPancreatic <- ggplot(rms:::Predict(coxPancreatic, HLIScore=seq(1, 20),
                                       ref.zero = TRUE,
                                        fun=exp),
                      xlab= "HLI Score",
                      ylab="HR",
                      adj.subtitle = FALSE) + 
                  geom_hline(yintercept=1, linetype="dashed") +
                  ggtitle("Linear")



pancreaticSplineOptions <- ggarrange(plotPancreatic5,
                                      plotPancreatic4,
                                      plotPancreatic,
                                      ncol=2, nrow=2,
                                     labels=c("a","b","c"))

pancreaticSplineOptions

```

\newpage

Kidney cancer


```{r kidney knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# kidney, 5 knots


plotKidney5 <- ggplot(rms:::Predict(coxspline5Kidney, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                      xlab= "HLI Score",
                      ylab="HR",
                      adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("5 knots")



# kidney, 4 knots



plotKidney4 <- ggplot(rms:::Predict(coxspline4Kidney, HLIScore=seq(1, 20),
                                    ref.zero = TRUE,
                                    fun=exp),
                      xlab= "HLI Score",
                      ylab="HR",
                      adj.subtitle = FALSE) + 
                geom_hline(yintercept=1, linetype="dashed") +
                ggtitle("4 knots")



#kidney, Linear



plotKidney <- ggplot(rms:::Predict(coxKidney, HLIScore=seq(1, 20),
                                   ref.zero = TRUE,
                                   fun=exp),
                        xlab= "HLI Score",
                        ylab="HR",
                        adj.subtitle = FALSE) + 
            geom_hline(yintercept=1, linetype="dashed") +
            ggtitle("Linear")



kidneySplineOptions <- ggarrange(plotKidney5,
                                 plotKidney4,
                                 plotKidney,
                                 ncol=2, nrow=2,
                                 labels=c("a","b","c"))

kidneySplineOptions

```


```{r render, eval=FALSE}

rmarkdown::render("C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/r_docs/rcs.Rmd", output_format= "pdf_document", output_file = "C:/Users/sch044/OneDrive - UiT Office 365/R/Paper1_new/output/rcs.pdf")

```




```{r p-linearity, include=FALSE}

 
library(lmtest)
 
#m2 <- cph(Surv(time,status) ~  rcs(albumin, 4), data=d)
 
#anova(m2)
 
#m <- cph(Surv(time, status) ~  albumin, data=d)

 
#p.val <- 1- pchisq( (m2$loglik[2]- m$loglik[2]), 2 )

 
#anova(coxspline3Breast)
pLinearityBreast <- 1- pchisq((coxspline3Breast$loglik[2] - coxBreast$loglik[2]), 2)
pLinearityBreast

 
#anova(coxspline3Breast, coxBreast)

anova(coxspline3Colorectal)

 
pLinearityColorectal <- 1- pchisq((coxspline3Colorectal$loglik[2] - coxColorectal$loglik[2]), 2)
pLinearityColorectal

 
anova(coxspline5Lung)

pLinearityLung <- 1- pchisq((coxspline5Lung$loglik[2] - coxLung$loglik[2]), 2)
pLinearityLung

 
anova(coxspline3Endometrial)

pLinearityEndometrial <- 1- pchisq((coxspline3Endometrial$loglik[2] - coxEndometrial$loglik[2]), 2)
pLinearityEndometrial

anova(coxspline3Pancreatic)
 
pLinearityPancreatic<- 1- pchisq((coxspline3Pancreatic$loglik[2] - coxPancreatic$loglik[2]), 2)
pLinearityPancreatic

 
anova(coxspline3Kidney)

pLinearityKidney<- 1- pchisq((coxspline3Kidney$loglik[2] - coxKidney$loglik[2]), 2)
pLinearityKidney

 
