---
title: "Adjusted flexible models"
subtitle: "Healthy Lifestyle Index and Risk for Common Cancers Modeled with Restricted Cubic Splines (RCS)"

author: "Sairah L.F. Chen"
date: "08/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rms)
library(survival)
library(tidyverse)
library(ggpubr)
library(cowplot)

casesIncluded <- read.csv2(file="../data/casesIncluded27082020.csv")
```

This supplementary document provides information on the use of RCS to model non-linear Linear associations between the HLI score and risk of common cancers. As included in the main article, the six most common cancer types in the Norwegian Women and Cancer (NOWAC) cohort will be explored.

## Best fit models

<span style="color:gray">Figure 1. Non-linear association of HLI score and risk for common cancers. Shaded region visualises 95% CI. Restricted cubic splines fitted with three knots set at HLI score 1, 13, and 20 for each cancer model </span>

```{r three knot plots , echo=FALSE, warning=FALSE}

ddist <- datadist(casesIncluded)
options(datadist="ddist")


# breast, 3 knots
coxspline3Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxsplineLinear3Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 3) + 
                                HLIScore +
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

plotBreast3 <- ggplot(rms:::Predict(coxspline3Breast, HLIScore=seq(1, 20),
                                    fun = exp),
                            xlab= "HLI Score",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer") +
  theme(legend.position = "none")

plotBreast3

#crc, 3 knots

coxspline3Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotColorectal3 <- ggplot(rms:::Predict(coxspline3Colorectal, HLIScore=seq(1, 20),
                                        fun=exp),
                            xlab= "HLI Score",
                          ylab="HR",
                          adj.subtitle = FALSE) + 
  ggtitle("Colorectal cancer")

plotColorectal3

#lung, 3 knots

coxspline3Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotLung3 <- ggplot(rms:::Predict(coxspline3Lung, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                    adj.subtitle = FALSE) + 
  ggtitle("Lung cancer")

plotLung3
# endometrial, 3 knots
coxspline3Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

plotEndometrial3 <- ggplot(rms:::Predict(coxspline3Endometrial, HLIScore=seq(1, 20),
                                    fun=exp),
                        xlab= "HLI Score",
                      ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Endometrial cancer")

plotEndometrial3

# pancreatic, 3 knots
coxspline3Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotPancreatic3 <- ggplot(rms:::Predict(coxspline3Pancreatic, HLIScore=seq(1, 20),
                                    fun=exp),
                        xlab= "HLI Score",
                      ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Pancreatic cancer")

plotPancreatic3

# kidney, 3 knots

coxspline3Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 3) + education, data=casesIncluded, x=TRUE, y=TRUE)

plotKidney3 <- ggplot(rms:::Predict(coxspline3Kidney, HLIScore=seq(1, 20),
                                        fun=exp),
                            xlab= "HLI Score",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Kidney cancer")

plotKidney3

rcs3Knots <- plot_grid(plotBreast3,
             plotColorectal3, 
             plotLung3,
             plotEndometrial3,
             plotPancreatic3,
             plotKidney3,
             ncol=3, nrow=2)
rcs3Knots

#rcs3KnotsLabeled <- add_sub(rcs3Knots, 
#                                       "Figure 1. Non-linear association of HLI score and risk for common cancers. \nShaded region visualises 95% CI. Restricted cubic splines fitted with three \nknots set at HLI score 1, 13, and 20 for each cancer model", 
#                            size=12,
#                            x = 0,
#                            h = 0)
#ggdraw(rcs3KnotsLabeled)


png("nonlinear_plots.png")
rcs3Knots
dev.off()


```


## Postmenopausal breast cancer sensitivity analysis 

<span style="color:gray"> Nonlinear associations of HLI score excluding each lifestyle component one-by-one and risk for postmenopausal breast cancer. Shaded region visualises 95% CI. Restricted cubic splines fitted with three knots set at HLI score 1, 13, and 16 for each cancer model </span>


```{r, echo=FALSE, warning=FALSE}
casesIncluded$HLInoSmoking <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoPA <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$smokingScore +
  casesIncluded$BMIScore

casesIncluded$HLInoBMI <- casesIncluded$dietScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$smokingScore

casesIncluded$HLInoDiet<- casesIncluded$smokingScore +
  casesIncluded$alcoholScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

casesIncluded$HLInoAlcohol <- casesIncluded$dietScore +
  casesIncluded$smokingScore +
  casesIncluded$physicalActivityScore +
  casesIncluded$BMIScore

breastNoSmoking <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoSmoking, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoSmoking <- ggplot(rms:::Predict(breastNoSmoking, HLInoSmoking=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer without smoking") +
  theme(legend.position = "none")

plotbreastNoSmoking


breastNoPA <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoPA, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoPA <- ggplot(rms:::Predict(breastNoPA, HLInoPA=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer without PA") +
  theme(legend.position = "none")

plotbreastNoPA


breastNoBMI <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoBMI, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoBMI <- ggplot(rms:::Predict(breastNoBMI, HLInoBMI=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer without BMI") +
  theme(legend.position = "none")

plotbreastNoBMI


breastNoDiet <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoDiet, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoDiet <- ggplot(rms:::Predict(breastNoDiet, HLInoDiet=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer without diet") +
  theme(legend.position = "none")

plotbreastNoDiet


breastNoAlcohol <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotbreastNoAlcohol <- ggplot(rms:::Predict(breastNoAlcohol, HLInoAlcohol=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Breast cancer without alcohol") +
  theme(legend.position = "none")

plotbreastNoAlcohol


```

## Lung cancer sensitivity analysis 

<span style="color:gray"> Nonlinear associations of HLI score excluding each lifestyle component one-by-one and risk for lung cancer risk. Shaded region visualises 95% CI. Restricted cubic splines fitted with three knots set at HLI score 1, 13, and 16 for each cancer model </span>

```{r, echo=FALSE, warning=FALSE}
lungNoSmoking <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoSmoking, 3) + 
                          education +
                         smokingScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoSmoking <- ggplot(rms:::Predict(lungNoSmoking, HLInoSmoking=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without smoking",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Lung cancer without smoking") +
  theme(legend.position = "none")

plotLungNoSmoking

lungNoPA <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoPA, 3) + 
                          education +
                         physicalActivityScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoPA <- ggplot(rms:::Predict(lungNoPA, HLInoPA=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without PA",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Lung cancer without PA") +
  theme(legend.position = "none")

plotLungNoPA

lungNoBMI <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoBMI, 3) + 
                          education +
                         BMIScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoBMI <- ggplot(rms:::Predict(lungNoBMI, HLInoBMI=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without BMI",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Lung cancer without BMI") +
  theme(legend.position = "none")

plotLungNoBMI


lungNoDiet <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoDiet, 3) + 
                          education +
                         dietScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoDiet <- ggplot(rms:::Predict(lungNoDiet, HLInoDiet=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without diet",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Lung cancer without diet") +
  theme(legend.position = "none")

plotLungNoDiet


lungNoAlcohol <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLInoAlcohol, 3) + 
                          education +
                         alcoholScore, data=casesIncluded, x=TRUE, y=TRUE)

plotLungNoAlcohol <- ggplot(rms:::Predict(lungNoAlcohol, HLInoAlcohol=seq(1, 16),
                                        fun=exp),
                            xlab= "HLI Score without alcohol",
                          ylab="HR",
                      adj.subtitle = FALSE) + 
  ggtitle("Lung cancer without alcohol") +
  theme(legend.position = "none")

plotLungNoAlcohol


```


```{r, echo=FALSE, warning=FALSE}

ddist <- datadist(casesIncluded)
options(datadist="ddist")
coxspline5Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 5) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Breast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ rcs(HLIScore, 4) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          familyHistBC +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxBreast <- cph(Surv(ageMenopause, ageExit, statusBreast) ~ HLIScore + 
                   education +
                   HRTStatus +
                   OCEverUse +
                   parityCat +
                   familyHistBC +
                   ageMenarche +
                   breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline5Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Colorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxColorectal <- cph(Surv(ageEntry, ageExit, statusColorectal) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)


coxspline5Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Lung <- cph(Surv(ageEntry, ageExit, statusLung) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxLung <- cph(Surv(ageEntry, ageExit, statusLung) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline5Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 5) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Endometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ rcs(HLIScore, 4) + 
                          education +
                          HRTStatus +
                          OCEverUse +
                          parityCat +
                          ageMenarche +
                          breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)



coxEndometrial <- cph(Surv(ageMenopause, ageExit, statusEndometrial) ~ HLIScore + 
                   education +
                   HRTStatus +
                   OCEverUse +
                   parityCat +
                   ageMenarche +
                   breastfeedingCat, data=casesIncluded, x=TRUE, y=TRUE)

coxspline5Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxspline4Pancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxPancreatic <- cph(Surv(ageEntry, ageExit, statusPancreatic) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)


coxspline5Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 5) + education, data=casesIncluded, x=TRUE, y=TRUE)


coxspline4Kidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ rcs(HLIScore, 4) + education, data=casesIncluded, x=TRUE, y=TRUE)

coxKidney <- cph(Surv(ageEntry, ageExit, statusKidney) ~ HLIScore + education, data=casesIncluded, x=TRUE, y=TRUE)






AIC <- cbind(c(AIC(coxspline5Breast),
               AIC(coxspline4Breast),
               AIC(coxspline3Breast),
               AIC(coxBreast)),
             c(AIC(coxspline5Colorectal),
               AIC(coxspline4Colorectal),
               AIC(coxspline3Colorectal),
               AIC(coxColorectal)),
             c(AIC(coxspline5Lung),
               AIC(coxspline4Lung),
               AIC(coxspline3Lung),
               AIC(coxLung)),
             c(AIC(coxspline5Endometrial),
               AIC(coxspline4Endometrial),
               AIC(coxspline3Endometrial),
               AIC(coxEndometrial)),
             c(AIC(coxspline5Pancreatic),
               AIC(coxspline4Pancreatic),
               AIC(coxspline3Pancreatic),
               AIC(coxPancreatic)),
             c(AIC(coxspline5Kidney), 
               AIC(coxspline4Kidney), 
               AIC(coxspline3Kidney), 
               AIC(coxKidney)))
rownames(AIC) <- c("5 knots",
                   "4 knots",
                   "3 knots (best fit)",
                   "Linear")
colnames(AIC) <- c("Breast",
                   "Colorectal",
                   "Lung",
                   "Endometrial",
                   "Pancreatic",
                   "Kidney")

knitr::kable(as.data.frame(AIC), digits = 2,
             caption = "Table 1. AIC values for Linear HLI score and RCS HLI score models among common cancers")
```


## Alternate spline models

### Breast cancer

<span style="color:gray">Figure 2. Associations between HLI and risk for breast cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r breast knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# breast, 5 knots


plotBreast5 <- ggplot(rms:::Predict(coxspline5Breast, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# breast, 4 knots



plotBreast4 <- ggplot(rms:::Predict(coxspline4Breast, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")




#breast, Linear



plotBreast <- ggplot(rms:::Predict(coxBreast, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



breastSplineOptions <- ggarrange(plotBreast5,
                               plotBreast4,
                               plotBreast,
                               ncol=2, nrow=2)

breastSplineOptions

```

### Colorectal cancer

<span style="color:gray">Figure 3. Associations between HLI and risk for colorectal cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r crc knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# crc, 5 knots


plotColorectal5 <- ggplot(rms:::Predict(coxspline5Colorectal, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# crc, 4 knots



plotColorectal4 <- ggplot(rms:::Predict(coxspline4Colorectal, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")



#crc, Linear



plotColorectal <- ggplot(rms:::Predict(coxColorectal, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



colorectalSplineOptions <- ggarrange(plotColorectal5,
                               plotColorectal4,
                               plotColorectal,
                               ncol=2, nrow=2)

colorectalSplineOptions

```

### Lung cancer

<span style="color:gray">Figure 4. Associations between HLI and risk for lung cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r lung knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# lung, 5 knots


plotLung5 <- ggplot(rms:::Predict(coxspline5Lung, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# lung, 4 knots



plotLung4 <- ggplot(rms:::Predict(coxspline4Lung, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")



#lung, Linear



plotLung <- ggplot(rms:::Predict(coxLung, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



lungSplineOptions <- ggarrange(plotLung5,
                               plotLung4,
                               plotLung,
                               ncol=2, nrow=2)

lungSplineOptions

```

### Endometrial cancer

<span style="color:gray">Figure 5. Associations between HLI and risk for endometrial cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r endometrial knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# endo, 5 knots


plotEndometrial5 <- ggplot(rms:::Predict(coxspline5Endometrial, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# endo, 4 knots



plotEndometrial4 <- ggplot(rms:::Predict(coxspline4Endometrial, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")



#endo, Linear



plotEndometrial <- ggplot(rms:::Predict(coxEndometrial, HLIScore=seq(1, 20),
                                        fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



endometrialSplineOptions <- ggarrange(plotEndometrial5,
                                      plotEndometrial4,
                                      plotEndometrial,
                               ncol=2, nrow=2)

endometrialSplineOptions

```

### Pancreatic cancer

<span style="color:gray">Figure 6. Associations between HLI and risk for pancreatic cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r pancreatic knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# lung, 5 knots


plotPancreatic5 <- ggplot(rms:::Predict(coxspline5Pancreatic, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# lung, 4 knots



plotPancreatic4 <- ggplot(rms:::Predict(coxspline4Pancreatic, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")



#lung, Linear



plotPancreatic <- ggplot(rms:::Predict(coxPancreatic, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



pancreaticSplineOptions <- ggarrange(plotPancreatic5,
                                      plotPancreatic4,
                                      plotPancreatic,
                               ncol=2, nrow=2)

pancreaticSplineOptions

```

### Kidney cancer

<span style="color:gray">Figure 7. Associations between HLI and risk for kidney cancer. Plots display the HLI modeled as an RCS variable with five and four knots, and as an unaltered Linear variable. Knots are set at percentile locations. Shaded regions visualise 95% CI. </span>

```{r kidney knot plots, echo=FALSE}
ddist <- datadist(casesIncluded)
options(datadist="ddist")

# lung, 5 knots


plotKidney5 <- ggplot(rms:::Predict(coxspline5Kidney, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("5 knots")



# lung, 4 knots



plotKidney4 <- ggplot(rms:::Predict(coxspline4Kidney, HLIScore=seq(1, 20),
                   fun=exp),
       xlab= "HLI Score",
     ylab="HR",
     adj.subtitle = FALSE) + 
  ggtitle("4 knots")



#lung, Linear



plotKidney <- ggplot(rms:::Predict(coxKidney, HLIScore=seq(1, 20),
                                  fun=exp),
                      xlab= "HLI Score",
                    ylab="HR",
                   adj.subtitle = FALSE) + 
  ggtitle("Linear")



kidneySplineOptions <- ggarrange(plotKidney5,
                                      plotKidney4,
                                      plotKidney,
                               ncol=2, nrow=2)

kidneySplineOptions

```

## Significance of the non-linear term

```{r p-linearity, echo=TRUE}

library(lmtest)
#m2 <- cph(Surv(time, status) ~  rcs(albumin, 4), data=d)
#anova(m2)
#m <- cph(Surv(time, status) ~  albumin, data=d)

#p.val <- 1- pchisq( (m2$loglik[2]- m$loglik[2]), 2 )

anova(coxspline3Breast)

pLinearityBreast <- 1- pchisq((coxspline3Breast$loglik[2] - coxBreast$loglik[2]), 2)
pLinearityBreast

pLinearityBreastOpposite <- 1- pchisq((coxBreast$loglik[2]) - coxspline3Breast$loglik[2], 2)
pLinearityBreastOpposite

lrtest(coxspline3Breast, coxBreast)

#anova(coxspline3Breast, coxBreast)
anova(coxspline3Colorectal)

pLinearityColorectal <- 1- pchisq((coxspline3Colorectal$loglik[2] - coxColorectal$loglik[2]), 2)
pLinearityColorectal


anova(coxspline3Lung)

pLinearityLung <- 1- pchisq((coxspline3Lung$loglik[2] - coxLung$loglik[2]), 2)
pLinearityLung


anova(coxspline3Endometrial)

pLinearityEndometrial <- 1- pchisq((coxspline3Endometrial$loglik[2] - coxEndometrial$loglik[2]), 2)
pLinearityEndometrial


anova(coxspline3Pancreatic)

pLinearityPancreatic<- 1- pchisq((coxspline3Pancreatic$loglik[2] - coxPancreatic$loglik[2]), 2)
pLinearityPancreatic


anova(coxspline3Kidney)

pLinearityKidney<- 1- pchisq((coxspline3Kidney$loglik[2] - coxKidney$loglik[2]), 2)
pLinearityKidney

``` 




